Enzo: An Adaptive Mesh Refinement Code for Astrophysics

### 2013-06-25 conclusions.tex
Section 14 edited (conclusions).

With the continual rapid development of computer hardware, it makes
sense to not only review \enzo's current capabilities, but to look
toward its future development in view of predicted technological
trends. These trends in supercomputing hardware suggest that
substantial modifications to \enzo's core infrastructure, and very
possibly some of the core algorithms, will be required. More
specifically, the progression involves the usage of specialized
large-core-count, vectorized computing units such as graphics
processing units or chips like the Intel Xeon Phi, as well as
precipitously decreasing amounts of RAM per computing core.  The
former trend means that the amount of processing power per compute
node will continue to increase, likely much faster than the bandwidth
between nodes, and will require tremendous reduction in inter-node
(and possibly inter-CPU) communication in order to maintain code
scalability.  Also, much of the current code will need to be rewritten
to take advantage of the vector nature of these CPUs, making
assumptions that are quite unlike those made in much of the current
codebase.  The latter trend means that duplication of data -- for
example, the grid hierarchy -- must be effectively eliminated to save
memory, and all inter-core and inter-node communication must be
carefully thought through to minimize the amount of data moved.  An
additional challenge as one goes to core counts in the tens to
hundreds of millions (or more) is that the reliability of individual
computing elements will become much more of an issue, requiring
robustness to hardware failure to be built into the code at some
level.  Furthermore, we are nearing the physical limits of transistor
speed and latency~\citep{feynman1999feynman}, meaning that simple
hardware improvements will not make these challenges disappear, and
careful thought (and the rewriting of a great deal of code) must take
place!  These challenges are not unique to the \enzo\ code, and in
fact are faced by effectively all applications that wish to take
advantage of new computational architectures. We therefore anticipate
that \enzo\ (or a code that has the capabilities of \enzo, from a
user's point of view) will continue to be usable at the
largest scales on such machines.

>>>>

With the continual rapid development of computer hardware, it makes
sense to not only review \enzo's current capabilities, but to look
toward its future development in view of predicted technological
trends. These trends in supercomputing hardware suggest that
substantial modifications to \enzo's core infrastructure, and very
possibly some of the core algorithms, will be required. More
specifically, the progression involves the usage of specialized
large-core-count, vectorized computing units such as graphics
processing units or chips like the Intel Xeon Phi, as well as
precipitously decreasing amounts of RAM per computing core.  The
former trend means that the amount of processing power per compute
node will continue to increase, likely much faster than the bandwidth
between nodes, and will require tremendous reduction in inter-node
(and possibly inter-CPU) communication in order to maintain code
scalability.  Also, much of the current code will need to be rewritten
to take advantage of the vector nature of these CPUs, making
assumptions that are quite unlike those made in much of the current
codebase.  The latter trend means that duplication of data -- for
example, the grid hierarchy -- must be effectively eliminated to save
memory, and all inter-core and inter-node communication must be
carefully thought through to minimize the amount of data moved.  An
additional challenge as one goes to core counts in the tens to
hundreds of millions (or more) is that the reliability of individual
computing elements will become much more of an issue, requiring
robustness to hardware failure to be built into the code at some
level.  Furthermore, we are nearing the physical limits of transistor
speed and interconnect latency~\citep{feynman1999feynman}, meaning that simple
hardware improvements will not make these challenges disappear, and
careful thought (and the rewriting of a great deal of code) must take
place!  These challenges are not unique to the \enzo\ code, and in
fact are faced by effectively all applications that wish to take
advantage of new computational architectures. We therefore anticipate
that \enzo\ (or a code that has the capabilities of \enzo, from a
user's point of view) will continue to be usable at the
largest scales on such machines.
